Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2021-02-17 numerics_paper/version2/numerical_chj.tex

\subsection{Rusanov's Riemann solver}\label{sec:rusanov}
The Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
upper bound on the wave speeds appearing in the true solution of the Riemann problem.
In this work, we take $\lambda_{i-1/2}^{\max}$ to be the upper bound
from \cite[Prop. 3.7]{azerad2017well}.
The waves are then given by
\begin{align}\label{rusanov_waves}
  \W_{i-1/2}^{1,\Rus}=\bfq_{i-1/2}-\bfq_{i-1}, \qquad
  \W_{i-1/2}^{2,\Rus}=\bfq_{i}-\bfq_{i-1/2},
\end{align}
where the intermediate state $\bfu_{i-1/2}$ is determined by imposing conservation:
\begin{align}\label{rus_cons}
  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfq_{i})-\bff(\bfq_{i-1}).
\end{align}
The fluctuations are given by \eqref{fluct} with $s_{i-1/2}^1=-\lambda_{i-1/2}^{\max}$
and $s_{i-1/2}^2=\lambda_{i-1/2}^{\max}$, and the first- and second-order methods based
on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.

@@

\subsection{The correction fluxes}
In this section we explain the choice of wave speeds \eqref{s_p}.
For the moment, we consider \eqref{blended-flux-1}-\eqref{lambda_p} without
entropy stabilization; i.e. we set $\lambda^{\min}_\iph=0$
for now.
We use $\W^{p,\Rus}$ to denote the waves in the Rusanov's
solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
The first-order method given by \eqref{first-order_via_fluct} with the numerical
flux \eqref{blended-flux-1} can also be written in
terms of fluctuations:
\begin{subequations}\label{ev_fluctuations}
\begin{align}
  \apdq_\imh & = \frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^{p,\Roe}, \\
  \amdq_\iph & = \frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^{p,\Roe}.
\end{align}
\end{subequations}
To implement the correction fluxes required for the second-order scheme \eqref{second-order_via_fluct},
we must also define a set of waves and corresponding speeds.
Using only the waves from the Roe solver, it is in general not possible
to choose speeds that yield the fluctuations \eqref{ev_fluctuations},
and we instead use \eqref{s_p}.
Using this in \eqref{correction-fluxes} and \eqref{ev_fluctuations} in 
\eqref{second-order_via_fluct}, we obtain (in the absence of limiting)
the second-order scheme
\begin{align}\label{LWL}
  \bfq_i^{n+1}=\bfq_i^n
  &-\frac{\Delta t}{2 \Delta x}
  \sum_p
  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}(\lambda_{i+1/2}^p)^2 \right)\W_{i+1/2}^{p,\Roe}
  +
  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^{p,\Roe}\right],
\end{align}
It is clear again that if $\theta_{i\pm 1/2}=0$ we recover the standard
second-order Lax-Wendroff method based on Roe's Riemann solver.
We now show that if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method
based on Rusanov's Riemann solver.

>>>>

\subsection{Rusanov's Riemann solver}\label{sec:rusanov}
Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
upper bound on the wave speeds appearing in the true solution of the Riemann problem.
In this work, we take $\lambda_{i-1/2}^{\max}$ to be the upper bound
from \cite[Prop. 3.7]{azerad2017well}.
The waves are then given by
\begin{align}\label{rusanov_waves}
  \W_{i-1/2}^{1,\Rus}=\bfq_{i-1/2}-\bfq_{i-1}, \qquad
  \W_{i-1/2}^{2,\Rus}=\bfq_{i}-\bfq_{i-1/2},
\end{align}
where the intermediate state $\bfu_{i-1/2}$ is determined by imposing conservation:
\begin{align}\label{rus_cons}
  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfq_{i})-\bff(\bfq_{i-1}).
\end{align}
The fluctuations are given by \eqref{fluct} with $s_{i-1/2}^1=-\lambda_{i-1/2}^{\max}$
and $s_{i-1/2}^2=\lambda_{i-1/2}^{\max}$, and the first- and second-order methods based
on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.

@@

\subsection{The correction fluxes}
In this section we explain the choice of wave speeds \eqref{s_p}.
For the moment, we consider \eqref{blended-flux-1}-\eqref{lambda_p} without
entropy stabilization; i.e. we set $\lambda^{\min}_\iph=0$
for now.
We use $\W^{p,\Rus}$ to denote the waves in the Rusanov
solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
The first-order method given by \eqref{first-order_via_fluct} with the numerical
flux \eqref{blended-flux-1} can also be written in
terms of fluctuations:
\begin{subequations}\label{ev_fluctuations}
\begin{align}
  \apdq_\imh & = \frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^{p,\Roe}, \\
  \amdq_\iph & = \frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^{p,\Roe}.
\end{align}
\end{subequations}
To implement the correction fluxes required for the second-order scheme \eqref{second-order_via_fluct},
we must also define a set of waves and corresponding speeds.
Using only the waves from the Roe solver, it is in general not possible
to choose speeds that yield the fluctuations \eqref{ev_fluctuations},
and we instead use \eqref{s_p}.
Using this in \eqref{correction-fluxes} and \eqref{ev_fluctuations} in 
\eqref{second-order_via_fluct}, we obtain (in the absence of limiting)
the second-order scheme
\begin{align}\label{LWL}
  \bfq_i^{n+1}=\bfq_i^n
  &-\frac{\Delta t}{2 \Delta x}
  \sum_p
  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}(\lambda_{i+1/2}^p)^2 \right)\W_{i+1/2}^{p,\Roe}
  +
  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^{p,\Roe}\right],
\end{align}
It is clear again that if $\theta_{i\pm 1/2}=0$ we recover the standard
second-order Lax-Wendroff method based on Roe's Riemann solver.
We now show that if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method
based on Rusanov's Riemann solver.
